### 2012-11-16 doc/progress_plan/report.tex
Added crf mailing list problem appendix

The language that Partridge will be written in is one of the most fundamental
and important choices to be made. The language choice effects not only the end
performance of the application, but the speed at which the project is developed
and the portability of the end code \cite{britton2008}. To avoid a learning
overhead, it was decided that the programming language should be one that the
author is familiar with. This reduced the choice of language down to C, Java
and Python.

The C programming language was invented by Kernighan and Ritchie and
published in 1978\cite{ritchie1978c}. The language is small and
optimised\cite{prinz2002c} and therefore compiled C programs tend to run
extremely fast. Unfortunately, C is designed to be general and lack restriction
\cite{ricthie1978c} which is often an advantage for programmers who favour
optimised code over excessive error checking. However, this means that
debugging C programms can be quite long winded and challenging. Since the NLP
aspects of Partridge present an amibitious challenge in themselves, having to
debug applications without help from a managed programming environment is not
desirable.

The Java programming language on the other hand, does provide memory management
and error checking\cite{Coffey2008} at the cost of program performance. Java is
a pseudo-compiled language that is translated into bytecode then interpreted at
runtime by a Virtual Machine(VM) on the client computer. Gosling (2000)
describes java as a ``general-purpose...  programming language [that allows]
developers to write a program once and then be able to run it everywhere on the
internet\cite{gosling2000java}." Java provides a stable programming environment
and error checking, both of which would be advantageous in the development of
Partridge. Java's syntax is very verbose and a lot of `boiler-plate' code must
be written before a program can be executed.  Therefore Java is not suitable as
a rapid prototyping language. Since Partridge's development requires a steep
learning curve for the author, a language that allows quick prototyping with
minimal programming.

Python is a programming language with a famously shallow learning curve,
readable syntax and dynamic inspection that facilitates rapid prototyping
\cite{bird2009natural}. The language was invented by Guido Van Rossum as a
multi-use, flexible, interpreted scripting language, extensible via compiled C
modules\cite{An93pythonfor}. Developing Partridge in python
would provide a flexible prototyping environment and readable syntax allowing
for investigative work to be carried out quickly and efficiently with minimal
overheads for writing boiler-plate code and debugging. The ability to extend
the language using native code would give Partridge an even greater advantage.
Since NLP techniques can be quite processor intensive and Python is an
interpreted language, and in its very nature, slower than a compiled language,
inefficient sections of Python code could be re-written as C extensions and
compiled into the application. For these reasons, Python was selected as the
preferred programming language for Partridge development.

@@

It is possible to use a separate software stack Such as Apache web server
(\url{http://httpd.apache.org/}) and PHP (\url{http://www.php.net}) as a web
presentation layer. However, this would involve adding more software to the
project and building an interface between the PHP scripts and Partridge's
python backend. To prevent unnecessary complication, this approach will be
avoided and instead, a Python web toolkit will be used and integrated directly
with the backend.

Django (\url{http://www.djangoproject.com/}) is a Python web framework with
integrated database management and templating capabilities \cite{django2012}.
The framework provides automatic management of all aspects of your
application's data storage systems unless explicitely disabled. The framework
also uses a fairly stricty Model-View-Controller pattern and this is enforced
in the structure of the application code. These features may be useful in a
general web application such as a news site. Partridge requires a much more
complex internal structure to take into account the machine learning and NLP
aspects of the system. Having to `shoehorn' Partridge into a specific structure
to make it compatible with Django adds extra overhead to the project and is not
desirable.

Flask (\url{http://flask.pocoo.org}) is a python `microframework' for web
development. Ronacher (2012) writes that ``Flask aims to keep the core simple
but extensible...[and] won't make any decisions for you\cite{flask2012}." It
uses a simple templating system that can optionally be swapped out for an
equivalent system and does not impose any limitations or requirements on the
layout of the program or the type of database that is used. Flask integrates
very simply into existing software using Python annotations\emph{(Ibid)}. This approach
means that the framework could be adapted to Partridge very quickly, rather
than requiring Partridge to be adapted to the framework. Therefore, Flask has
been chosen as the preferred web framework for use in the project.

\subsubsection{Natural Language Libraries} \label{sec:libschoice}.  There are
several existing libraries to facilitate Natural Language Processing.  Many are
written for Java \cite{mallet2002}\cite{cunningham2011text} and are very
complex or not well documented. The Natural Language Toolkit (NLTK) is a simple
and intuitive library written for Python. Bird(2009) states that NLTK was
designed ``to provide an intuitive framework along with substantial building
blocks, giving users a practical knowledge of NLP\cite{bird2009natural}". The
project is relatively mature in comparison to the above named Java libraries.
There is also a free book that accompanies the project available at
\url{http://nltk.org/book/} which provides a huge amount of information on how
to implement many popular NLP techniques using the library. For this reason
NLTK was chosen as the primary accompanying library for the project.

@@

@@

The Python version of SAPIENTA was then installed

@@

>>>>

The language that Partridge will be written in is one of the most fundamental
and important choices to be made. The language choice effects not only the end
performance of the application, but the speed at which the project is developed
and the portability of the end code \cite{britton2008}. It was decided that
Python would be the best language for developing Partridge.

Python is a programming language with a famously shallow learning curve,
readable syntax and dynamic inspection that facilitates rapid prototyping
\cite{bird2009natural}. The language was invented by Guido Van Rossum as a
multi-use, flexible, interpreted scripting language, extensible via compiled C
modules\cite{An93pythonfor}. Developing Partridge in python
would provide a flexible prototyping environment and readable syntax allowing
for investigative work to be carried out quickly and efficiently with minimal
overheads for writing boiler-plate code and debugging. The ability to extend
the language using native code would give Partridge an even greater advantage.
Since NLP techniques can be quite processor intensive and Python is an
interpreted language, and in its very nature, slower than a compiled language,
inefficient sections of Python code could be re-written as C extensions and
compiled into the application.

Other languages that were considered were Java and C. However, it was decided
that Python's clean syntax and ease of use for prototyping were preferable to
the C and Java style syntax. As stated above, C may still be used in part for
developing Python extensions to improve the overall performance of the
application.

@@

For this purpose, the Flask (\url{http://flask.pocoo.org}) library was chosen.
Flask is a python `microframework' for web development. Ronacher (2012) writes
that ``Flask aims to keep the core simple but extensible...[and] won't make any
decisions for you\cite{flask2012}." It uses a simple templating system that can
optionally be swapped out for an equivalent system and does not impose any
limitations or requirements on the layout of the program or the type of
database that is used. Flask also integrates very simply into existing software
using Python annotations\emph{(Ibid)}.

An alternative approach that was considered was using a separate software
stack such as the Apache web server (\url{http://httpd.apache.org/}) and PHP 
(\url{http://www.php.net}). However, this would involve adding more software to the
project and building an interface between the PHP scripts and Partridge's
python backend. The Django framework (\url{http://www.djangoproject.com/}) for Python was also
considered but it was determined to be overly complicated in terms of
Partridge's requirements.

\subsubsection{Natural Language Libraries} \label{sec:libschoice}.

It was decided that a pre-existing NLP library should be used to provide
implementations of existing algorithms and techniques for Partridge. For this
purpose the The Natural Language Toolkit (NLTK) was selected. The NLTK is a
simple and intuitive library written for Python. Bird(2009) states that NLTK
was designed ``to provide an intuitive framework along with substantial
building blocks, giving users a practical knowledge of
NLP\cite{bird2009natural}". NLTK provides implementations for sentence
splitting, tagging (the process of analysing sentences and classifying words as
verbs, nouns etc.), and information retrieval from text \emph{(Ibid)}. NLTK
also provides implementations of classifiers and a framework for extracting
features to be used for classification from text. There is also a free book
that accompanies the project available at \url{http://nltk.org/book/} which
provides a huge amount of information on how to implement many popular NLP
techniques using the library.

Some other libraries were considered\cite{mallet2002}\cite{cunningham2011text}.
However, these were written for Java and Partridge is a python project. They
also require a steeper learning curve than NLTK.

@@

The `mega-journals' PLOS ONE \url{http://www.plosone.org/} and arXiv
\url{http://www.arxiv.org/} were suggested as sources for more open access
articles that could be added to Partridge. Both of these sites were found to
contain large volumes of open access papers. However, most of these papers were
published as PDF files which meant that some conversion would be required to
make them compatible with Partridge.

@@

The Python version of SAPIENTA was then downloaded and a test executed.
Unfortunately there were several data files missing from the package that had
to be acquired from Dr Liakata.

SAPIENTA for Python also relies upon a package called CRFSuite which implements
Conditional Random Fields, a method for segmenting and labelling sequence
data\cite{CRFsuite}. This library did not compile properly on the test
environment and its creator had to be contacted via a mailing list (See
Appendix \ref{sec:crfletter}.

@@

\section{Email for CRFSuite}
\label{sec:crfemail}
\input{crfletter}
